---
title: "Homework 5"
author: "Gianni Spiga"
date: '2022-10-27'
output: html_document
---

```{r, message = FALSE, warning=FALSE}
library(survival)
library(KMsurv)
library(ggplot2)
library(ggpubr)
library(survminer)
library(plotly)
library(muhaz)
library(ggthemes)
```

```{r}
vars <- c("id", "clinic", "status", "time", "prison", "methadone")
addicts <-
  read.table(
    url(
      "https://dmrocke.ucdavis.edu/Class/BST222.2022.Fall/addicts.txt"
    ),
    header = F,
    col.names = vars
  )
#addicts
```

## Question 1

```{r}
#change variables to factors to be used in Cox PH
addicts$clinic <-
  factor(addicts$clinic, labels = c("Clinic1", "Clinic2"))
addicts$prison <- factor(addicts$prison, labels = c("No", "Yes"))

#head(addicts)

surv_object <- Surv(time = addicts$time, event = addicts$status)
addicts.cox.strat <-
  coxph(surv_object ~ strata(clinic) + prison + methadone, data = addicts)
summary(addicts.cox.strat)
```

## Question 2

```{r}
addicts.zph.stra <- cox.zph(addicts.cox.strat)
addicts.zph.stra

#For clinic
ggcoxzph(addicts.zph.stra[1], se = FALSE, font.main = 12, ggtheme = theme_solarized(), point.col = "#0096FF", ylim = c(-4,4))

#For prison
ggcoxzph(addicts.zph.stra[2], se = FALSE, font.main = 12, ggtheme = theme_solarized(), point.col = "#0096FF", ylim = c(-0.3,0.3))
```
We can see from the plot that the prison variable has a sinusoidal pattern of variation. Meanwhile, our methadone variable seems to have little evidence of a pattern, with a plot that shows random scatter. 


## Question 3


```{r, warning = FALSE}
addicts.mart <- residuals(addicts.cox.strat, type = "martingale")

addicts.cs <- addicts$status - addicts.mart

#Cumulative hazard 
surv.csr <- survfit(Surv(addicts.cs, addicts$status) ~1, type = "fleming-harrington", data = addicts)

#plot(surv.csr, fun ="cumhaz")


cumHazPlot <-
  ggsurvplot(
    surv.csr,
    fun = "cumhaz",
    conf.int = TRUE,
    palette = c("#581845"),
    ggtheme = theme_solarized()
  ) + ggtitle("Cumulative Hazard for Cox-Snell Residuals")
#ggplotly(cumHazPlot[[1]])


ggplotly(cumHazPlot$plot + geom_abline())
```
We can see that after time about 2.5, our cumulative hazard does not follow our straight trend line. This gives us the impression that our model lacks in being a good fit for the data. 

## Question 4

```{r}
#This time, without methadone
addicts.coxNoMeth <- coxph(surv_object ~ strata(clinic) + prison, data = addicts)
addicts.mart <- residuals(addicts.coxNoMeth, type = "martingale")


lowessOBJ <- as.data.frame(lowess(addicts$methadone, addicts.mart))

ggplotly(
  ggplot() + aes(addicts$methadone, addicts.mart) + geom_point(col = "#FFA000") + labs(x = "Methadone", y = "Martingale Residuals", title = "Martingale Residuals vs. Methadone Dosage") + geom_line(data = lowessOBJ, aes(x = x, y = y), col = "#388E3C") + theme_solarized()
)
```

Since our lowess line seems pretty linear, we would advise that a transformation of the data is not neccessary. 

## Question 5

```{r}
# With Linear predictor
addicts.predict <- predict(addicts.cox.strat)
addicts.mart <- residuals(addicts.cox.strat, type = "martingale")

#Deviance and df beta
addicts.dev <- residuals(addicts.cox.strat, type = "deviance")
addicts.dfb <- residuals(addicts.cox.strat, type = "dfbeta")

#MArtingale vs Linear Predictor
ggplotly(
  ggplot() + aes(
    x = addicts.predict,
    y = addicts.mart,
    label = rownames(addicts)
  ) + geom_point() + labs(x = "Linear Predictor", y = "Martingale Residual", title = "Martingale Residuals vs Linear Predictor") + theme_solarized()
)

# Deviance vs Linear Predictor
ggplotly(
  ggplot() + aes(
    x = addicts.predict,
    y = addicts.dev,
    label = rownames(addicts)
  ) + geom_point() + labs(x = "Linear Predictor", y = "Deviance Residual", title = "Deviance Residuals vs Linear Predictor") + theme_solarized()
)

# DfBeta vs Linear Predictor
# Clinic
# ggplotly(
#   ggplot() + aes(x = as.numeric(rownames(addicts)), y = addicts.dfb[, 1]) + geom_point() + labs(x = "Observation Number", y = "dfbeta for Clinic Type", title = "dfbeta Values for Observation Number by Clinic Type") + theme_solarized()
# )

# Prison
ggplotly(
  ggplot() + aes(x = as.numeric(rownames(addicts)), y = addicts.dfb[, 1]) + geom_point() + labs(x = "Observation Number", y = "dfbeta for Prison Status", title = "dfbeta Values for Observation Number by Prison Status") + theme_solarized()
)

#Methadone
ggplotly(
  ggplot() + aes(x = as.numeric(rownames(addicts)), y = addicts.dfb[, 2]) + geom_point() + labs(x = "Observation Number", y = "dfbeta for Methadone Level", title = "dfbeta Values for Observation Number by Methadone Level") + theme_solarized()
)


```
### Observations to Examine by Residuals and Influence 

* Martingale Residuals
  + 8,9, 12, 26, 54
* Deviance Residuals
  + 8, 123, 156, 175
* Prison Influence
  + 8, 9 ,12, 26
* Methadone Level Influence
  + 26, 70, 84, 156
  
  From the residual plots, the most important pbservations to examine would be 8,9 26, 70, and 156. Let's observe these two observations:
  
```{r}
addicts[c(8, 9, 26, 70, 156),]
``` 

Observations 8 and 26 were prisoners who were censored observations in our data. 70 was a censored observation who was not a prisoner, but had a very long time. Given their long survival times, they could have felt they no longer needed to be part of the study. These prisoners also had low methadone levels. Observations 9 and 156 were recorded as leaving the clinic. 156, had a very high methadone level however. 