---
title: "Homework 4"
author: "Gianni Spiga"
date: "2022-10-21"
output: html_document
---

```{r, message = FALSE, warning=FALSE}
library(survival)
library(KMsurv)
library(ggplot2)
library(ggpubr)
library(survminer)
library(plotly)
library(muhaz)
library(ggthemes)
```

```{r}
vars <- c("id", "clinic", "status", "time", "prison", "methadone")
addicts <-
  read.table(
    url(
      "https://dmrocke.ucdavis.edu/Class/BST222.2022.Fall/addicts.txt"
    ),
    header = F,
    col.names = vars
  )
#addicts
```

## Question 1

```{r}
#change variables to factors to be used in Cox PH
addicts$clinic <- factor(addicts$clinic,labels=c("Clinic1","Clinic2"))
addicts$prison <- factor(addicts$prison,labels=c("No","Yes"))

head(addicts)

surv_object <- Surv(time = addicts$time, event = addicts$status)
addicts.cox <- coxph(surv_object ~ clinic + prison + methadone, data = addicts)
#summary(addicts.cox)

addicts.zph <- cox.zph(addicts.cox)
addicts.zph

#For clinic
ggcoxzph(addicts.zph[1], se = FALSE, font.main = 12, ggtheme = theme_solarized(), point.col = "#0096FF", ylim = c(-7,7))

#For prison
ggcoxzph(addicts.zph[2], se = FALSE, font.main = 12, ggtheme = theme_solarized(), point.col = "#0096FF", ylim = c(-4,4))

#For methadone/
ggcoxzph(addicts.zph[3], se = FALSE, font.main = 12, ggtheme = theme_solarized(), point.col = "#0096FF", ylim = c(-0.5,0.5))
```

When we look at the plots of our Schoenfield Residuals, as well as our p-values, we can see that both methadone and prison variables are significantly proportional while our clinic variable violates our proportinality assumption at any reasonable $\alpha$.

## Question 2

```{r, warning = FALSE}
addicts.mart <- residuals(addicts.cox, type = "martingale")

addicts.cs <- addicts$status - addicts.mart

#Cumulative hazard 
surv.csr <- survfit(Surv(addicts.cs, addicts$status) ~1, type = "fleming-harrington", data = addicts)

#plot(surv.csr, fun ="cumhaz")


cumHazPlot <-
  ggsurvplot(
    surv.csr,
    fun = "cumhaz",
    conf.int = TRUE,
    palette = c("#581845"),
    ggtheme = theme_solarized()
  ) + ggtitle("Cumulative Hazard for Cox-Snell Residuals")
#ggplotly(cumHazPlot[[1]])


ggplotly(cumHazPlot$plot + geom_abline())
```

We can see that since our Cox-Snell residuals curve fits a standard line quite well, we can say we do not observe a lack of fit. 

## Question 3

```{r}
#This time, without methadone
addicts.coxNoMeth <- coxph(surv_object ~ clinic + prison, data = addicts)
addicts.mart <- residuals(addicts.cox, type = "martingale")

# With Linear predictor
addicts.predict <- predict(addicts.cox)

lowessOBJ <- as.data.frame(lowess(addicts$methadone, addicts.mart))

ggplotly(
  ggplot() + aes(addicts$methadone, addicts.mart) + geom_point(col = "#FFA000") + labs(x = "Methadone", y = "Martingale Residuals", title = "Martingale Residuals vs. Methadone Dosage") + geom_line(data = lowessOBJ, aes(x = x, y = y), col = "#388E3C") + theme_solarized()
)


#Create categorical methadone
addicts$catMeth <-
  cut(addicts$methadone,
      c(0, 60, 79, 200),
      labels = c("<60", "60-79", "80+"))

addicts$catMeth <- as.factor(addicts$catMeth)

#New cox model
addicts.coxCatMeth <-
  coxph(surv_object ~ clinic + prison + catMeth, data = addicts)


#We print AIC using drop1
# This is with methadone as quantitative
drop1(addicts.cox, test = "Chisq")

#This is with methadone as categorical 
drop1(addicts.coxCatMeth, test = "Chisq")
```

We can see that the model with methadone as a quantitative has a better AIC than the one with methadone as categorical. Along with that, the quantitative methadone model has has a smaller p-value for the ChiSq test, providing greater statistically significant strentgh. We can conclude that replacing the quantitavie methadone with a 3-category factor does matter, as it hurts the strength of our model. 

## Question 4

```{r}



```